Modeling Galaxy Spectral Energy Distributions in the SDSS Spectroscopic Survey

### 2002-02-08 docs/paper/paper.tex
call maggies maggies
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@21 75dff457-8751-0410-ab1a-cd701654ce6b
docs
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@28 75dff457-8751-0410-ab1a-cd701654ce6b

\shortauthors{Blanton {\it et al.} (2000)}
\shorttitle{Low Redshift Galaxy Evolution}

@@

%% Mark off your abstract in the ``abstract'' environment. In the manuscript
%% style, abstract will output a Received/Accepted line after the
%% title and affiliation information. No date will appear since the author
%% does not have this information. The dates will be filled in by the
%% editorial office after submission.
\begin{abstract}
We present a method of inferring galaxy spectral energy distributions
(SEDs) from photometric observations of galaxies in broad band
filters. Using the Sloan Digital Sky Survey (SDSS) observations in the
optical wavelength regime, we demonstrate that the method yields
robust results. We show how the method can be used to recover galaxy
magnitudes in fixed frame bandpasses over a range of redshifts. We
compare our method to others, including galaxy spectrophotometry. For
the SDSS, fixed frame magnitudes around the SDSS $r$, $i$, and $z$
bands are highly reliable, at the level of a few percent; $g$-band
fixed frame magnitudes are somewhat more uncertain, and $u$-band fixed
frame magnitudes are highly uncertain, at the level of 10--20\%. All
code and templates from this paper are public.
\end{abstract}

@@

In the future, observations of galaxies (and indeed of any
astronomical sources) will be performed using integral field
spectrographs or equivalent devices which combine high spatial
resolution, high spectral resolution, and high signal-to-noise. By
that time, the phenomena of broad-band filters and the quaint
terminology surrounding their usage --- magnitudes, $K$-corrections,
color terms, {\it etc.} --- will have long since been
forgotten. However, today, virtually all observations of galaxies are
made through such filters, and special care has to be taken to recover
knowledge of galaxy spectral energy distributions (SEDs) from these
observations. This paper describes how to do so in a way well-suited
for recovering (for example) fixed-frame magnitudes of objects over a
large range of redshifts.

Photometric surveys of the sky observe galaxies in bandpasses which
are fixed in the rest-frame of the observer. However, spectra of
distant galaxies are redshifted due to the expansion of the
universe. Thus, the more distant the galaxy, the further the observed
bandpass is blueshifted relative to the rest-frame spectral energy
distribution (SED) of the object observed. To compare observations of
galaxies at different redshifts, it is important to account for the
fact that one is observing different parts of the galaxy SED.

In the past, this accounting has in general come in the form of
``$K$-corrections'' applied to the observed magnitudes. That is, a
function $K(z)$ is added to the standard cosmological bolometric
distance modulus $\mathrm{DM}(z)$. Sometimes a single function has
been applied regardless of the galaxy type, though more recently it
has become standard to use a discrete set of $K(z)$ functions
depending on galaxy type (based either on morphology or spectral
features). As described in \cite{oke68a} and later papers, the
function is based on the projection of an assumed galaxy SED
redshifted to $z$ onto the filter response.

@@

Our approach here is to develop a method for inferring the underlying
SEDs of a set of galaxies at a range of redshifts by requiring that
their SEDs all be drawn from a similar population. For each galaxy, we
will recover a model SED, which can be used to synthesize the galaxy's
magnitude in any bandpass. In addition, this method provides a natural
way to infer an estimate of the flux within the whole optical range.
The approach is equivalent to the photometric metric redshift
estimation methods of \citet{csabai99a} and \citet{budavari00a}, 
except we use slightly different coordinate systems.

Section \ref{sedfit} describes our method of fitting galaxy SEDs to
broad-band photometry. An implementation of the method in C and IDL is
provided at a URL on the World-Wide Web, through a public CVS
repository, or by request.  Section \ref{data} applies the method to
galaxies in the SDSS, showing that the fits are robust. Section
\ref{kcorrection} shows how one derives $K$-corrections from the
results, and discusses how best to use the results of such
fits. Section \ref{conclusion} concludes and discusses future
development of the method described here.

@@

The magnitudes are calibrated $AB$, meaning that the system is
designed such that:
\begin{eqnarray}
m_{AB} &=& -48.60 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \lambda f(\lambda) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right]
\cr
&=& -48.60 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right],
\cr
\end{eqnarray}
where $R(\lambda)$ is the fraction of photons which are detected as a
function of wavelength (a unitless quantity).

>>>>

\shortauthors{Blanton {\it et al.} (2000)}
\shorttitle{Modeling Galaxy SEDs}

@@

%% Mark off your abstract in the ``abstract'' environment. In the manuscript
%% style, abstract will output a Received/Accepted line after the
%% title and affiliation information. No date will appear since the author
%% does not have this information. The dates will be filled in by the
%% editorial office after submission.
\begin{abstract}
Certain astronomical questions, such as the evolution of the
luminosity function of galaxies and the relationship between galaxy
luminosities and galaxy colors, require transforming inherently
redshift-dependent measurements of galaxy magnitudes into measures of
galaxy properties which are consistent over a range of redshifts.  As
part of the task of addressing these questions, which we will
investigate in subsequent papers, we implement here the method of
\citet{csabai00a} for fitting spectral energy distributions (SEDs) to
broad-band photometric observations, in the context of the optical
observations of the Sloan Digital Sky Survey (SDSS). Linear
combinations of four spectral templates can reproduce the five SDSS
magnitudes of galaxies to the precision of the photometry. Expressed
in the correct coordinate system, the distribution of the coefficients
of the templates is planar, and in fact nearly linear. The resulting
reconstructed SEDs (the sum of the templates) can be used to recover
fixed-frame magnitudes over a range of redshifts. We show that the
results of doing so yield consistent results within the SDSS Main
Galaxy Sample and the SDSS Luminous Red Galaxy Sample, in the sense
that within each sample the trend with redshift of intrinsic colors
for similar type galaxies are small. A potential problem in the
measurement of the SDSS filter response curves is noted. We also
validate the results using galaxy spectrophotometry.  All code and
templates from this paper are publicly available and well-documented;
they are designed such that they can easily be incorporated into other
applications. The code is easily adapted to handle galaxy observations
in any AB magnitude system in the rest-frame optical regime.
\end{abstract}

@@

In the future, we expect that observations of galaxies (and indeed of
any astronomical sources) will be performed using integral field
spectrographs or equivalent devices which combine high spatial
resolution and high spectral resolution. By that time, the phenomena
of broad-band filters and the quaint terminology surrounding their
usage --- magnitudes, $K$-corrections, color terms, {\it etc.} ---
will have long since been forgotten. However, today, virtually all
observations of galaxies are made through such filters, and special
care has to be taken to recover knowledge of galaxy spectral energy
distributions (SEDs) from these observations. This paper focuses on a
method of facing this problem which (in the optical wavelength regime)
is fast, robust, and consistent over a large range of redshifts.

This task is nontrivial because SEDs of galaxies contain important
information on scales considerably smaller than the width of typical
broad-band filters. It is further complicated by the fact that SEDs of
distant galaxies are redshifts due to the expansion of the
universe. Thus, the more distant the galaxy, the further the observed
bandpass is blueshifted relative to the rest-frame spectral energy
distribution of the object observed.

In the past, accounting for these effects has in general come in the
form of ``$K$-corrections'' applied to the observed magnitudes. That
is, a function $K(z)$ is added to the standard cosmological bolometric
distance modulus $\mathrm{DM}(z)$. Sometimes a single function has
been applied regardless of the galaxy type, though more recently it
has become standard to use a discrete set of $K(z)$ functions
depending on galaxy type (based either on morphology or spectral
features). As described in \cite{oke68a} and later papers, the
function is based on the projection of an assumed galaxy SED
redshifted to $z$ onto the measured instrument response as a function
of wavelength (including the effects of the atmosphere, the
reflectivity of the mirrors, the filter transmission, and the response
of the CCD device or photographic emulsion).

@@

Our approach here is to use a method for inferring the underlying SEDs
of a set of galaxies at a range of redshifts by requiring that their
SEDs all be drawn from a similar population. For each galaxy, we will
recover a model SED, which can be used to synthesize the galaxy's
magnitude in any bandpass. In addition, this method provides a natural
way to infer an estimate of the flux within the whole optical range.
The approach is equivalent (nearly identical) to the photometric
metric redshift estimation methods of \citet{csabai99a} and
\citet{budavari00a}, except we use slightly different coordinate
systems.

We implemented this system and are publishing it in order to prepare
the reader for upcoming papers which will rely heavily on the
reliability of the SED fits described here.  However, having
implemented the fits, there is no reason not to share them with the
community; thus, this paper serves the secondary purpose of describing
a piece of software being released to the astronomical community. The
computer software created is a C shared object library, around which
we have written both stand-alone C programs and IDL routines. We have
made the source code available publicly, through a URL on the World
Wide Web and through a public CVS repository. Improvements or ports to
other languages implemented by users may be incorporated into the code
upon request. The conditions of use for the code are that this paper
is cited in any resulting refereed journal article and that the
version of the code used is specified in any such paper. The version
of the code used to make the figures for this paper is {\tt kcorrect
v1_2}.

Section \ref{sedfit} describes our method of fitting galaxy SEDs to
broad-band photometry. Section \ref{data} applies the method to
galaxies in the SDSS, showing that the fits are robust. Section
\ref{kcorrection} shows how one derives $K$-corrections from the
results, and discusses how best to use the results of such
fits. Section \ref{conclusion} concludes and discusses future
development of the method described here.

@@

The magnitudes are calibrated $AB$, meaning that the system is
designed such that:
\begin{eqnarray}
m_{AB} &=& -48.60 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \lambda f(\lambda) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right]
\cr
&=& -2.41 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right],
\cr
\end{eqnarray}
where $R(\lambda)$ is the fraction of photons which are detected as a
function of wavelength (a unitless quantity). The difference in the
zeropoint is simply the factor of speed of light $c$ (expressed in
\AA\ s$^{-1}$) between the frequency and wavelength scales. {\bf
clearer on units}
